---
title: "chensus"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{chensus}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup, warning=FALSE, message=FALSE}
library(chensus)
library(dplyr)
```

# Introduction

The `chensus` package estimates population frequencies, means,
proportions and confidence intervals from surveys conducted by
the Federal Statistical Office (FSO):

-   structural survey: *Strukturerhebung* (SE) / *relevé structurel* (RS),
-   mobility and transport survey: *Mikrozensus Mobilität und Verkehr*
    (MZMV) / *Microrecensement mobilité et transports* (MRMT).

## Example Data

In this vignette, we demonstrate the main features of the package using the built-in `nhanes` dataset, which contains a subset of data from the [National Health and Nutrition Examination Survey](https://wwwn.cdc.gov/Nchs/Nhanes/2015-2016/DEMO_I.htm) for the period 2015-2016 (more with `?nhanes`). Its structure is similar to FSO survey data in that it contains `strata` and `weights` columns and demographic features such as `gender` and `household_size`. 
## Notes on Example Data and Interpretation

The `nhanes` dataset included in this package is used solely as a publicly available example to demonstrate how to use `chensus`. In practice, the package is designed for use with FSO's structural survey and mobility and transport survey data. For any other use, results are illustrative only.

The `nhanes` dataset is included with the package. Let's load and inspect it:
```{r}
data("nhanes")
dplyr::glimpse(nhanes)
nhanes %>% count(gender)
nhanes %>% count(household_size)
```
# Structural Survery Example Analysis

Suppose we want to estimate the population in the `nhanes` dataset by gender and age. We can use the main analysis function `chensus::se_total()`:

```{r}
se_total(
  data = nhanes,
  weight = "weights",
  strata = "strata",
  condition = c("gender", "age")
)
```
If on the other hand we would like to estimate the average household size then we would use the function `chensus::se_mean_num()`:

```{r}
se_mean_num(
  data = nhanes, 
  variable = household_size, 
  strata = strata,
  weight = weights)
```

We could also estimate the proportion of males and females in the `nhanes` survey, `gender` being a categorical variable. 

```{r}
se_mean_cat(
  data = nhanes,
  variable = "gender",
  weight = "weights",
  strata = "strata"
)
```


